\chapter{DTEAM: Pure Process Intelligence}

In previous iterations, DTEAM spent considerable computational effort parsing noisy XES event logs, handling missing data, and reconstructing a plausible model of past events. It sat downstream of the execution.

\section{From Reconstruction to Verified State}

DTEAM is no longer responsible for discovering truth from operational exhaust. In the \textbf{uni*} ecosystem, truth has already been canonicalized, admitted, and receipted before it reaches the intelligence layer. DTEAM consumes only verified canonical facts and receipt-backed projections. It does not depend on, inspect, or know the private \texttt{unibit} representation.

\section{Autonomic Discovery over Lawful Geometry}

DTEAM acts as an autonomic observer inside the system, analyzing the geometric space. It applies continuous reinforcement learning over verified facts to steer future executions.

\section{Search, RL, and Routing over Coordinate-State Space}

When DTEAM proposes route or process intent, it uses search and reinforcement learning algorithms over the lawful geometry. Because the world is already mapped to the $64^3$ operational depth, DTEAM proposes geometric trajectories. These intents are then handed to MuStar to be compiled into kinetic motion packets, completing the operational loop.

\section{Nanosecond-Scale Operational Intelligence}

With the operational globe pinned inside the CPU's $L_1$ cache (utilizing the 32 KiB footprint for both TruthBlock and Scratchpad), the \texttt{unibit} substrate achieves over 4.3 billion cryptographically verified state transitions per second. This absolute utilization of zero-heap, branchless execution unlocks unprecedented capabilities for DTEAM:

\subsection{Exhaustive Future Simulation via Reinforcement Learning}
Because the operational globe is mathematically bounded to exactly 262,144 discrete memory positions, the intelligence layer does not guess or wait for downstream databases. At billions of operations per second, DTEAM can execute continuous reinforcement learning and instantaneous search algorithms over the entire global coordinate-state space. It simulates millions of "what-if" supply chain scenarios to mathematically prove the optimal path before executing a physical move.

\subsection{Instantaneous Planetary Rerouting Under Chaos}
Real-world chaos---such as port closures or embargos---traditionally causes delays while traversing complex graph databases. In this framework, a localized crisis simply flips a "law mask" or "prereq mask" at a specific geometric cell. Operating at maximum CPU utilization, DTEAM instantly re-evaluates the geometric constraints of every entity globally, calculates new trajectories, and verifies their liveness and deadlock-freedom in a fraction of a second.

\subsection{Real-Time, Verifiable Executive Projections}
Under a maximized \texttt{unibit} substrate, the engine inherently generates cryptographically secure BLAKE3 receipts and exact sparse bit-deltas for every state transition. Feeding this verified operational physics directly into specialized visualization engines guarantees an unbroken, mathematically proven model of the supply chain, entirely eliminating the possibility of phantom data.